Analysis and Figures for Gallego et al, 2020

### 2020-09-25 Manuscript/What.lies.ahead.reviewed.tex
almost ready for submission

Our sampled areas differed in the environmental variables driving changes in carbonate chemistry. San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1C); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there. By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supporting Information).

@@

Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with \textit{insect} v1.1 \cite{insect}; where \textit{insect} could not place individual taxa, we supplemented assignment by classification against a custom COI database using \textit{anacapa} \cite{curd2019anacapa} and \textit{bowtie2} \cite{langmead2012fast}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time (``eDNA index'' \cite{kelly2019understanding}), using pooled data from technical replicates and mean proportions across biological replicates. We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions. In the Supplementary information, we provide a function in R to calculate this index.

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole. We fit a normal distribution to our 2017 environmental observations to create baseline conditions, and modeled the change in mean parameter values between 2017 and 2095 as a linear function of time. We then used the modeled distribution of environmental parameters to generate 1000 simulations of each year scenario. The scenario labeled as 2095 is the set of parameters falling within the 95\% percentile in simulations for the years 2091-2095. See the R code in the supplementary information (lines 98-135).

To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supporting Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (dark gold) and San Juan Island (turquoise). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay. Boxplots represent the variability in richness across all time points for samples taken at the indicated site. \textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately; note change in the scale of the y-axis. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as vertical facets.}}
\label{fig:Fig1}
\end{figure}
%%%%%

%% All changes due to new run of data are shown with *** before them for ease of search
Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs). Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs. San Juan Island; F = 1.6184; p $<$ 0.01). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig. 2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature. In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp. C are more prevalent.  The community common in the high temperature range (cluster 3 in Fig. 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species. Cluster 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material). On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Likely winners under future conditions are more widely scattered among higher taxonomic groups. The haptophytes \textit{Emiliania huxleyi} and \textit{Chrysochromulina} sp. the dinoflagellate \textit{Alexandrium} sp. all find more suitable habitat in both of our study areas. Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas. See Supporting Information for a complete list of taxon-specific projections.

@@

Among the taxa surveyed, diatoms are of particular interest for their ubiquity in the world's oceans and their important roles in marine food webs \cite{armbrust2009life, wasmund2017diatom}, as well as in ecological and evolutionary theory \cite{margalef11978}. Our model suggests that diatoms will decrease in richness between the present and 2095, particularly in the Hood Canal, where extreme temperatures are more common. Although the most prevalent response among diatoms is a decrease in suitability, some substantial variability in responses exists within the group. For example, the centric diatom \textit{Coscinodiscus} spp., which is a food source for \textit{Acartia} spp. copepods \cite{jansen2008copepods} and many other animal species, will see future suitable habitat only in colder waters such as those in San Juan Island, while \textit{Skeletonema} spp. and the harmful algal bloom (HAB)-forming species \textit{Pseudo-nitzschia} spp. will see their habitat suitability remain constant or slightly increased, especially at low pH levels (see Supplemental Material).

@@

 We are grateful to the UW Center for Environmental Genomics and the Conservation Biology Molecular Genetics Laboratory at NOAA Fisheries (Montlake) for access to lab equipment and associated expertise and support. Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.

>>>>

Our sampled areas differed in the environmental variables driving changes in carbonate chemistry. San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1C); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there. By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supplementary Information).

@@

Code for all quality-screening and bioinformatics is available in the Supplementary Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with \textit{insect} v1.1 \cite{insect}; where \textit{insect} could not place individual taxa, we supplemented assignment by classification against a custom COI database using \textit{anacapa} \cite{curd2019anacapa} and \textit{bowtie2} \cite{langmead2012fast}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supplementary Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time (``eDNA index'' \cite{kelly2019understanding}), using pooled data from technical replicates and mean proportions across biological replicates. We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions. In the Supplementary Information, we provide a function in R to calculate this index.

@@

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole. We fit a normal distribution to our 2017 environmental observations to create baseline conditions, and modeled the change in mean parameter values between 2017 and 2095 as a linear function of time. We then used the modeled distribution of environmental parameters to generate 1000 simulations of each year scenario. The scenario labeled as 2095 is the set of parameters falling within the 95\% percentile in simulations for the years 2091-2095. See the R code in the Supplementary Information (lines 98-135).

To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supplementary Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.

@@

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (dark gold) and San Juan Island (turquoise). \textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay. Boxplots represent the variability in richness across all time points for samples taken at the indicated site. \textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately; note change in the scale of the y-axis. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU).}}
\label{fig:Fig1}
\end{figure}
%%%%%

Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions. Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than 50.8M sequences from 778 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control the dataset included $\sim${}45M sequences, from 4849 unique amplicon sequence variants (ASVs). Of these, 1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supplementary Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs. San Juan Island; F = 1.6184; p $<$ 0.01). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supplementary Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the community linked with colder water and higher-pH (community 2 in Fig. 2C) showed higher eDNA indexes of diatoms (\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature. In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp. C are more prevalent.  The community common in the high temperature range (cluster 3 in Fig. 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species. Cluster 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplementary Information). On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature). \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supplementary Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

@@

Likely winners under future conditions are more widely scattered among higher taxonomic groups. The haptophytes \textit{Emiliania huxleyi} and \textit{Chrysochromulina} sp. the dinoflagellate \textit{Alexandrium} sp. all find more suitable habitat in both of our study areas. Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas. See Supplementary Information for a complete list of taxon-specific projections.

@@

Among the taxa surveyed, diatoms are of particular interest for their ubiquity in the world's oceans and their important roles in marine food webs \cite{armbrust2009life, wasmund2017diatom}, as well as in ecological and evolutionary theory \cite{margalef11978}. Our model suggests that diatoms will decrease in richness between the present and 2095, particularly in the Hood Canal, where extreme temperatures are more common. Although the most prevalent response among diatoms is a decrease in suitability, some substantial variability in responses exists within the group. For example, the centric diatom \textit{Coscinodiscus} spp., which is a food source for \textit{Acartia} spp. copepods \cite{jansen2008copepods} and many other animal species, will see future suitable habitat only in colder waters such as those in San Juan Island, while \textit{Skeletonema} spp. and the harmful algal bloom (HAB)-forming species \textit{Pseudo-nitzschia} spp. will see their habitat suitability remain constant or slightly increased, especially at low pH levels (see Supplementary Information).

@@

 We are grateful to the UW Center for Environmental Genomics and the Conservation Biology Molecular Genetics Laboratory at NOAA Fisheries (Montlake) for access to lab equipment and associated expertise and support. Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment. The comments of two reviewers helped us improve the manuscripts and encourage us to streamline the analysis of our data.

### 2020-09-25 Scripts/Manuscript.Main.Analysis.Rmd
almost ready for submission

Because the standarization was done for each region separately, we have to convert them into standard values seapratedly as well.

@@

Make the ridges plot

```{r envRidges}
data.plankton %>% 
  select(sample = event, Area, TA = mean.TA, DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
  distinct() %>%  
  filter(pH > 7.4) %>% 
  ungroup() %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F, sep = "_" ) %>% 
  pivot_longer(cols = c(-Area, -Site, -Date,  -sample), names_to =   "Predictor", values_to  = "value")  %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) %>% 
  mutate(Month = month(ymd(Date), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_density_ridges_gradient() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal", axis.title.x = element_blank()) -> ridges

@@

THe top will be the environmental gradients, the lower panels will have the map and the boxplot. Create a custom palette of colors for the two regions

@@

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  cap.functions.only3things(Area) -> joined.dataset.for.CAP
```

@@

joined.dataset.for.CAP %>% pull(CAP.significance)
```

@@

  CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {

    fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")

  } ),

@@

Clusters.analysis.and.cap %>% 
  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>%right_join(.y) ),
         full.cluster.info = map2(full.cluster.info,CAP.plot, ~ .y$sites %>% 
                                    as.data.frame() %>% 
                                    rownames_to_column("event") %>% 
                                    right_join(.x) %>% 
                                    mutate(pH = round(pH,1)))

@@

  })) %>% pull(Hand.made.CAP) -> leftplot.Fig2

 leftplot.Fig2 %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top

@@

ggsave(plot,filename = here("Figures","Supplementary",paste0("Supp.info.San.Juan.Island.Fig2", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina")

@@

### Richness by phyla with time
 First prepare the grid - expand it to include the taxa
```{r}
full.grid %>% 
  select(Area,  Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) -> full.grid

@@

full.grid %>% 
  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
  summarise(richness = sum(present)) %>% 
#  left_join(richness.by.phylum) %>% 
  group_by(phylum) %>% 
  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
```

```{r plot4 function}
plot4 <- function(tibble, response) {

@@

```{r plot4 in play}
richness.by.phylum %>% 
  nest(-phylum) %>% 
  mutate(plot = map(data, ~ plot4(.x, response = st.richness)))  -> list.of.richness.plots

@@

Following reviewer's suggestions, we are limiting the information displayed in the main Figures to two of most important phyla for coastal ecosystems dinamics: Diatoms and Dinoflagellates.
```{r}
layout <- "
AA
BB
##
CC
DD
##
EE
GG
HH
"
wrap_plots(A = left.plot[["Dinophyceae"]] + labs(title = "A.1", tag = "Dinophyceae") ,
           B = left.plot.top[["Dinophyceae"]] + guides(fill = "none")+ labs(title  ="A.2"),
          # F = plot_spacer(),
           C = left.plot[["Bacillariophyta"]] + labs(title ="B.1", tag = "Bacillariophyta"),
           D = left.plot.top[["Bacillariophyta"]] + guides(fill = "none")+ labs(title ="B.2"),
           E = left.plot[["Haptophyceae"]] + labs(title ="C.1", tag = "Haptophyceae"),
           G = left.plot.top[["Haptophyceae"]] + guides(fill = "none")+ labs(title  ="C.2"),
           H = guide_area(), design = layout, heights = c(3,2,1,3,2,1,3,2,1), guides = "collect") &
  theme(plot.tag.position = c(0.1,1),
        plot.tag = element_text(face = "bold")) -> left

@@

@@

richness.by.phylum.by.point %>%
  filter(phylum == phyl) %>% 
  ggplot(aes())+
  geom_histogram(bins = 100, aes(x =  richness, group = Year, color = Year, after_stat(ndensity)),fill = "white", position = "identity") +
  geom_vline(data = medians, aes(xintercept = grp.median, color = Year), linetype = "dashed") +
  facet_wrap (~Area)

@@

Now the Global - richness by point

>>>>

Because the standarization was done for each region separately, we have to convert them into standard values separately as well.

@@

Make the box plots

```{r envRidges}
data.plankton %>% 
  select(sample = event, Area, TA = mean.TA, DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
  distinct() %>%  
  filter(pH > 7.4) %>% 
  ungroup() %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F, sep = "_" ) %>% 
  pivot_longer(cols = c(-Area, -Site, -Date,  -sample), names_to =   "Predictor", values_to  = "value")  %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) %>% 
  mutate(Month = month(ymd(Date), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_boxplot() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal", axis.title.x = element_blank()) -> ridges

@@

The top will have the map and the boxplots, the lower panel will be the environmental gradients. Create a custom palette of colors for the two regions

@@

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_", remove = F) %>% 
  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
  cap.functions.only3things(Area) -> joined.dataset.for.CAP
```

@@

@@

  # CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {
  # 
  #   final <- kmeans(.x$sites, .y, nstart = 15)
  # 
  #   fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")
  # 
  # } ),

@@

Clusters.analysis.and.cap %>% 
  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>% right_join(.y) ),
         full.cluster.info = map2(full.cluster.info,CAP.plot, ~ .y$sites %>% 
                                    as.data.frame() %>% 
                                    rownames_to_column("event") %>% 
                                    right_join(.x) %>% 
                                    mutate(pH = round(pH,1)))

@@

  })) %>% pull(Hand.made.CAP) -> left_top

 left_top %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top

@@

ggsave(plot,filename = here("Figures","Supplementary",paste0("Supp.info.San.Juan.Island.Fig2", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina") 
plot
```

@@

### Richness by phyla across the pH and Temperature continuum

THe idea of this panel is to account for the limitations of a simulation of future scenarios, and just look at the behaviour of the model in the continuum - we know that temperature and pH will change in a direction, and we can explore broad patterns in the raster grid.

 First prepare the grid - expand it to include the taxa information.
```{r}
full.grid %>% 
  select(Area,  Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) -> full.grid

@@

full.grid %>% 
  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
  summarise(richness = sum(present)) %>% 
  group_by(phylum) %>% 
  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
```

Create a function to plot the richness across temperature and pH
```{r plot4 function}
plot4 <- function(tibble, response) {

@@

Create a richness plot for each phylum, pull them and name them
```{r plot4 in play}
richness.by.phylum %>% 
  nest(-phylum) %>% 
  mutate(plot = map(data, ~ plot4(.x, response = st.richness)))  -> list.of.richness.plots

@@

Following reviewer's suggestions, we are limiting the information displayed in the main Figures to three important phyla for coastal ecosystems dynamics: Diatoms, Haptophytes and Dinoflagellates.
```{r}
layout <- "
AA
BB
##
CC
DD
##
EE
GG
HH
"
wrap_plots(A = left.plot[["Dinophyceae"]] + labs(title = "A.1", tag = "Dinophyceae") ,
           B = left.plot.top[["Dinophyceae"]] + guides(fill = "none")+ labs(title  ="A.2"),
          # F = plot_spacer(),
           C = left.plot[["Bacillariophyta"]] + labs(title ="B.1", tag = "Bacillariophyta"),
           D = left.plot.top[["Bacillariophyta"]] + guides(fill = "none")+ labs(title ="B.2"),
           E = left.plot[["Haptophyceae"]] + labs(title ="C.1", tag = "Haptophyceae"),
           G = left.plot.top[["Haptophyceae"]] + guides(fill = "none")+ labs(title  ="C.2"),
           H = guide_area(), design = layout, heights = c(3,2,1,3,2,1,3,2,1), guides = "collect") &
  theme(plot.tag.position = c(0.1,1),
        plot.tag = element_text(face = "bold")) -> left

@@

You can find similar figures for all phyla found in the Folder `Figures/Supplemental`

@@

richness.by.phylum.by.point %>%
  filter(phylum == phyl) %>% 
  ggplot(aes())+
  geom_histogram(bins = 100, aes(x =  richness, group = Year, color = Year, after_stat(ndensity)),fill = "white", position = "identity") +
  geom_vline(data = medians, aes(xintercept = grp.median, color = Year), linetype = "dashed") +
  facet_wrap (~Area)+
  ggtitle(phyl)

@@

Now the Global - richness by point. If you want to see the change on any other phyla, type(`explore.phylum("phylum.name")`), you can see here plots for Dinoflagellates and diatoms
